# Preliminaries and Set-Up

R is an open-source programming language for statistical computing that allows you to carry out a wide range of data analysis tasks. One of the big advantages of using R is that it has a very large user community among social scientists and statisticians, who frequently publish packages (which one may think of as workbooks of sorts, containing a well-integrated set of R functions, scripts, data, and documentation) to accomplish certain tasks or implement given procedures. These packages are then shared with the broader community; at this point, anyone who needs to accomplish the tasks to which the package addresses itself can use the package in the context of their own projects. The ability to use published packages considerably simplifies the work of applied social scientists using R; it means that they rarely have to write code entirely from scratch, and can build on the code that others have published. This allows applied researchers to focus on substantive problems, without having to get too bogged down in complicated programming tasks. 

## Install Packages

In the context of this tutorial, generating maps of ICPSR data in R would be relatively complex if we had to write all our code from scratch. However, because we are able to make use of mapping and visualization packages written by other researchers, the task is considerably simpler, and will not require any complicated programming. 

In this tutorial, we will use a few packages to accomplish our goal of mapping tabular ICPSR data. They are:

+ *tmap*: The *tmap* package will allow us to create and customize a publishable map. 
+ *sf*: The *sf* package allows us to work with spatially explicit data within R.
+ *rnaturalearth*, *rnaturalearthdata*, *rgeos*: In order to visualize our data on a world map, we need a dataset of the world's country boundaries. One way to get such a dataset is to download it from a public repository, and then load it into R Studio. However, these packages allow us to load a spatial dataset of world boundaries into R Studio without having to actually download anything, which effectively saves us a few steps in the workflow. 
+ *readxl*: The *readxl* package (part of a broader suite of data science packages known as the "tidyverse") will allow us to load the tabular ICPSR dataset, which is provided as an Excel file on the ICPSR landing page, into R Studio. 
+ *dplyr*: The *dplyr* package (also a part of the tidyverse) will allow us to efficiently transform the structure of the initial ICPSR dataset into a more tractable form that is conducive to mapping. 

To install a package in R, we can use the ```install.packages``` function. A function is essentially a programming construct that takes a specified input, runs this input (called an "argument") through a set of procedures, and returns an output. In the code block below, the name of the package we want to install (here, "tmap") is enclosed within quotation marks and placed within parentheses after calling the ```install.packages``` function:

```{r, eval=FALSE}
## Installs "tmap" package
install.packages("tmap")
```

After running that line of code, the "tmap" function will effectively be downloaded to your computer. Now, let's install the other packages we mentioned above, using the same procedure: 

```{r, eval=FALSE}
## Installs remainder of necessary packages to complete exercise
install.packages("sf")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("rgeos")
install.packages("readxl")
install.packages("dplyr")
```

All of the packages we need are now installed!

## Load Libraries 

However, while our packages are installed in our R environment, they are not yet ready to use. Before we can use our packages, we must load them into our environment. You can think of the process of loading installed packages into a current R environment as analgous to opening up an application on your phone or computer after it has been installed (even if an application is installed, you can't use it until you open it!). To load (i.e. "open") an R package, we pass the name of the package we want to load as an argument to the ```library``` function. For example, if we want to load the "tmap" package into the current environment, we can type:

```{r, results=FALSE, message=FALSE}
## Loads "tmap" package
library(tmap)
```

At this point, the full suite of the "sf" package's functionality is available for us to use. 

Now, let's go ahead and load the remainder of the packages that we'll need:

```{r, results=FALSE, message=FALSE}
## Loads remainder of packages
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(tidyverse)
library(readxl)
library(dplyr)
```

At this point, the packages are loaded and ready to go! One important thing to note regarding the installation and loading of packages is that you only have to install packages once; after a package is installed, there is no need to subsequently reinstall it. However, you must load the packages you need (using the ```library``` function) every time you open a new R session. In other words, if you were to close R Studio at this point and open it up later, you would **not** need to install these packages again (Step 1a), but you **would** need to load the packages again (Step 1b). 

## Set Working Directory 

Before we can bring in our data, we have to specify the location on our computer where the data that we want to import is saved. This step is known as setting one's working directory. Before setting your working directory, make sure you've downloaded the ICPSR dataset, and have placed it in a directory (i.e. folder) on your computer that was specifically created to store this data. 

If you're unfamiliar with the concept of file paths, the easiest way to set your working directory is through the R Studio menu. To do so, follow these steps: 

+ First, Click on the "Session" button on the R Studio menu at the top of your screen, and then scroll down to the "Set Working Directory" button in the dialog box that opens up. 
+ When you hover over this button, a subsidiary dialog box that contains a button that says "Choose Directory" will open; click this "Choose Directory" button. 
+ In the dialog box that opens up, navigate to the directory that contains the ICPSR data, select it, and click "Open". At this point, your working directory should be set!

**Alternatively**, if you are familiar with the concept of file paths, and know the file path to the folder containing the downloaded ICPSR dataset, you can set the working directly using the ```setwd()``` function, where the argument to the function is the relevant file path enclosed in quotation marks. For example:

```{r}
## Sets working directory 
setwd("/Users/adityaranganath/Documents/git_repositories/icpsr_mapping_manual/tutorial_data")
```

