# Practice Exercises

In this section you'll have the opportunity to apply what you learned in Section 4, and practice some of your skills. The following three exercises are arrayed in order of increasing difficulty. 

## Practice Exercise 1

Modify the map we made in Section 4 in the following ways:

* Change the color scheme. What rationale informed your choice of this color scheme? 
* Change the legend breaks. Why did you chose the legend breaks that you did? 
* Replace your name with the current name in the "Map Credits" section. 

What spatial patterns do you notice in this map? Are there any patterns of clustering or dispersion that you find interesting? Do you notice anything surprising or unexpected? And, what future research questions might this map suggest? 

## Practice Exercise 2

Use the same dataset we used in Section 4, but this time please make an interactive map of the country-level average of the "Rigidity_Public_Health" variable. 

What spatial patterns do you notice in this map? Are there any patterns of clustering or dispersion that you find interesting? Do you notice anything surprising or unexpected? And, what future research questions might this map suggest? 

## Practice Exercise 3

Refer to the following [dataset](https://www.openicpsr.org/openicpsr/project/121001/version/V1/view?path=/openicpsr/121001/fcr:versions/V1/Index_IPR_Agriculture.xlsx&type=file), entitled "Worldwide Index of IPRs [Intellectual Property Rights] in Agriculture, 1961-2018", which is archived on ICPSR and authored by Mercedes Campi amd Alessandro Nuvolari.

Download the dataset, and make a map that shows country-level variation in the "Total_Index" variable in the year 2018. 

## Sample Code

The sample code in this section assumes you've already loaded all of the packages discussed in the tutorial, and have correctly set your working directory to the directory in which the dataset(s) are located. 

### Sample Code for Exercise 2

```{r, include=FALSE}
library(tmap)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(tidyverse)
library(readxl)
library(dplyr)
```

```{r, message=FALSE, warning=FALSE, echo=c(-1,-2)}
tmap_mode("plot")
setwd("/Users/adityaranganath/Documents/git_repositories/icpsr_mapping_manual/tutorial_data")

# Import "Dataset" sheet from ICPSR Excel file into R Studio, and assigns the dataset to an object called "covid_data"
covid_data<-read_excel("Gov_Responses2Covid19_last.xlsx", sheet="Dataset")

# Brings spatial dataset of country boundaries into R environment using the rnaturalearth package and assigns it to object named "country_boundaries"
country_boundaries<-ne_countries(scale="medium", returnclass="sf")

# Deletes Antarctica from "country_boundaries" object
country_boundaries<-country_boundaries %>% filter(iso_a3 !="ATA")

# Selects "iso" and "Rigidity_Public_Health" variables from ICPSR Dataset, and assigns this new dataset to object named "covid_data_public_health" 
covid_data_public_health<-covid_data %>% 
  select(iso, Rigidity_Public_Health) 

# Changes class of "Rigidity_Public_Health" field from character to numeric
covid_data_public_health$Rigidity_Public_Health<-as.numeric(covid_data_public_health$Rigidity_Public_Health)

# Calculate Country-Level Averages for "Rigidity_Public_Health" Index, and then assign this new dataset of country-level averages to a new object called "covid_data_PH_avg". 
covid_data_PH_avg<-covid_data_public_health %>% group_by(iso) %>% 
                                            summarize(mean_public_health=mean(Rigidity_Public_Health, na.rm=FALSE))

# Join dataset with country-level means for "Rigidity_Public_Health" index ("covid_data_PH_avg" ) to spatial dataset of world boundaries ("country_boundaries"), based on common 3-Digit ISO Codes; then relocate "name" and "mean_economic" fields to the beginning of the dataset so that they're displayed on the map
worldmap_covid_data_PH<-full_join(country_boundaries, covid_data_PH_avg, by=c("iso_a3"="iso")) %>% 
                                relocate(name, mean_public_health)

# Make map and assigns to object called "covid_PH_webmap"
covid_PH_webmap<-tm_shape(worldmap_covid_data_PH)+
                            tm_polygons(col="mean_public_health", n=5, style="jenks", palette="Oranges")

# Set tmap mode to "view
tmap_mode("view")

# display map object in view mode
covid_PH_webmap
```

### Sample Code for Exercise 3

```{r, message=FALSE, warning=FALSE, echo=c(-1,-2)}
tmap_mode("plot")
setwd("/Users/adityaranganath/Documents/git_repositories/icpsr_mapping_manual/tutorial_data")

# Load Tabular Data and assign to object named "IPR_data" 
IPR_data<-read_excel("Index_IPR_Agriculture.xlsx", sheet="Total_Index")
View(IPR_data)

# Load Spatial Data and assign to object named "country_boundaries" 
country_boundaries<-ne_countries(scale="medium", returnclass="sf")
View(country_boundaries)

# Delete Antarctica from "country_boundaries" 
country_boundaries<-country_boundaries %>% filter(iso_a3 !="ATA")

# Check class of Index field in "IPR_data" and confirm it's numeric
class(IPR_data$`Total Index`)

# Select 2018 Observations from "IPR_data" and assign to new object named "IPR_data_2018"
IPR_data_2018<-IPR_data %>% filter(year==2018)

# View "IPR_data_2018" in R Studio Data Viewer
View(IPR_data_2018) 

# Join "Datasets"IPR_data-2018" to "country_boundaries" based on common ISO codes, and then assign the product of the join to a new object named "spatial_IPR_2018."
spatial_IPR_2018<-full_join(country_boundaries, IPR_data_2018, by=c("iso_a3"="Country_code")) %>% 
                    relocate(Country, 'Total Index')
View(spatial_IPR_2018)

# Make map of variation in agricultural IPR index in 2018, and assign to new object named "IPR_2018_map"
IPR_2018_map<-tm_shape(spatial_IPR_2018)+
                tm_polygons(col="Total Index",
                            n=6, 
                            style="jenks", 
                            palette="PuBu",
                            title="Agriculture IPR\nIndex, 2018")+
                tm_layout(frame=FALSE,
                          legend.outside=TRUE,
                          legend.outside.position=c("bottom"),
                          main.title = "Strength of Intellectual Property Protection in Agriculture, 2018",
                          main.title.position = "Center",
                          main.title.size=1,
                          inner.margins=c(0.06,0.10,0.10,0.08),
                          attr.outside=TRUE)+
                tm_credits("Map Author: YOUR NAME\nDataset Author: Mercedes Campi\nCRS:WGS84\nSource:ICPSR",
                           position=c(0.78,0),
                           size=0.38)
# Open "IPR_2018_map" 

IPR_2018_map

```



